---
title: "AnalysisBundeswahl2009"
author: "Duc-Anh Nguyen"
date: "2025-02-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read the dataset and check its features
```{r dataset}
wahl <- read.table("Wahl.txt",header = TRUE)
head(wahl)
skimr::skim(wahl)
```

## fit a logit model with all combinations of interactions, as there are 10 independent variables we will have
```{r }
# fit a logit-model with interactions
logit_interaction_model <- glm(WA~.*., family = binomial(), data = wahl)
summary(logit_interaction_model)
```
Not many stat. significant effects. I want to check for colinearity now.


## check for colinearity
We check for colinearity by checking for singularity in the design matrix
```{r }
X <- model.matrix(logit_interaction_model)
Matrix::rankMatrix(X)
dim(X)
```

We see that there are 56 columns but the rank is only 53, which means there are two effects that are having colinearity. Let us then do variables selection with BIC
```{r }
logit_BIC_model <- MASS::stepAIC(object = logit_interaction_model, direction = "both", k = log(nrow(wahl)), trace = FALSE)
summary(logit_BIC_model)
```
There are only 7 main effects and 2 interaction effects left, one for sex:living(west or east) and one for age:religion(not catholic). 

For main effects: age, being catholic, and following other religions are having a negative correlation with decision of voting for SPD instead of CCU. 

The interaction of sex:living in the west and age:being religious are showing negative effects on the chance of voting for SPD instead of CCU


Now we will consider the non-linear effect of Alter_z (age). We fit the age effect and the interaction of age:sex now non-linearly with P-splines with max. number of nodes of k = 30
```{r }
logit_GAM_model <- mgcv::gam(formula = WA~s(Alter_z, bs = "ps", k = 30) + s(Alter_z, bs = "ps", k = 30, by = Geschlecht) + Geschlecht + West + Gewerk + Demozufrieden + Religion_kath  + Geschlecht*West, family = binomial(), data = wahl)

summary(logit_GAM_model)

# visualising the P-splines
par(mfrow = c(2,1))
plot(x = logit_GAM_model, shade = TRUE, select = 1, cex.main = 0.8, cex = 0.8, cex.lab = 0.8, cex.axis = 0.8, main = "Glatter Effekt des Alters für Frauen")
plot(x = logit_GAM_model, shade = TRUE, select = 2, cex.main = 0.8, cex = 0.8, cex.lab = 0.8, cex.axis = 0.8, main = "Glatter Effekt des Alters für Männer")
```
Attention: Although the lower graphic shows a linear effect for male voters, we need to add the effect of both plots in order to interpret the varying effect for males while simutaniously considering effect of age 


# Now for evaluation with ROC
```{r }
library(tidyverse)
library(ggplot2)
# Logit-Model:
wahl$BIC_predict <-
  predict(object = logit_BIC_model, type = "response")

# plot1 <-
#   ggplot(mapping = aes(m = BIC_predict, d = WA), data = wahl) + geom_roc() + geom_abline(intercept = 0, slope = 1) + ylab("Sensitivität") + xlab("1- Spezifität") + theme + ggtitle("Logit-Modell")

wahl$GAM_predict <-
  predict(object = logit_GAM_model, type = "response")

# plot2 <-
#   ggplot(mapping = aes(m = GAM_predict, d = WA), data = wahl) +
#   geom_roc(n.cuts = 0) + geom_abline(intercept = 0, slope = 1) +
#   ylab("Sensitivität") + xlab("1- Spezifität") + theme +
#   ggtitle("Logit-GAM")

 # Gemeinsame Visualisierung
# grid.arrange(plot1, plot2, ncol = 2)
```


getting the AUC
```{r }
library(pROC)
# logit_model
auc_logit_BIC <- roc(response = as.factor(wahl$WA), predictor = wahl$BIC_predict)$auc
auc_logit_GAM <- roc(response = as.factor(wahl$WA), predictor = wahl$GAM_predict)$auc

auc_logit_BIC
auc_logit_GAM
```
# now let us try to predict the chance of:
+ an catholic woman in her 25 in union from east germany, who is satisfied with democracy system in germany choosing SPD
+ a jewish man in his 75 not in union from west germany, who is not satisfied with democracy system 

```{r }


newdata <-
  data.frame(
    Alter_z = c(-25.5, 24.5),
    Geschlecht = c(0, 1),
    West = c(0, 1),
    Gewerk = c(1, 0),
    Demozufrieden = c(1, 0),
    Religion_kath = c(1, 0),
    Religion_Rest = c(0, 1)
  )
newdata 

# glm
predict(object = logit_BIC_model, newdata = newdata, type = "response")
# gam
mgcv::predict.gam(object = logit_GAM_model, newdata = newdata, type = "response")
```

